noam_data_exploration_v1

### 2024-08-16 noam_data_exploration_v3.Rmd
Fixed site level synchrony and portfolio effects

site_level <- transect_level %>%
  dplyr::group_by(year, island_side, site, habitat, taxa) %>%
  dplyr::summarise(mean_site_cover = mean(mean_cover), 
            site_se = sd(mean_cover)/sqrt(n()))
```

@@

# portfolio and synchrony need to be calculated at quad level, so taking the mean of the quad level for these
synchrony_portfolio_site <- diversity_stability_quadrat %>%
  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
  dplyr::summarise(
            algae_loreau_mean = mean(loreau_synchrony),
            algae_loreau_se = sd(loreau_synchrony)/sqrt(n()),
            algae_portfolio_mean = mean(portfolio_effect),
            algae_portfolio_se = sd(portfolio_effect)/sqrt(n()))

# merge together
diversity_stability_site <- merge(diversity_cv_site, synchrony_portfolio_site, by = c("site", "habitat", "island_side", "site_habitat"))
# write_csv(diversity_stability_site, here("data", "algae_diversity_stability_metrics_site.csv"))

>>>>

site_level <- long_data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
  dplyr::summarise(mean_site_cover = mean(percent_cover), 
            site_se = sd(percent_cover)/sqrt(n()))

@@

# portfolio and synchrony need to be calculated at quad level, so taking the mean of the quad level for these
# synchrony_portfolio_site <- diversity_stability_quadrat %>%
#   dplyr::group_by(site, habitat, island_side, site_habitat) %>%
#   dplyr::summarise(
#             algae_loreau_mean = mean(loreau_synchrony),
#             algae_loreau_se = sd(loreau_synchrony)/sqrt(n()),
#             algae_portfolio_mean = mean(portfolio_effect),
#             algae_portfolio_se = sd(portfolio_effect)/sqrt(n()))
# NOT DOING IT LIKE THIS ANYMORE! DOING IT BY RECALCULATING SYNCHRONY WITH SITE AS REPLICATE
lm_synchrony_site <- codyn::synchrony(df = site_level,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "mean_site_cover",
                                   metric = "Loreau",
                                   replicate.var = "site_habitat") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_site) = c("site_habitat", "loreau_synchrony") # rename cols for merge later

wide_site_level <- data %>%
  dplyr::group_by(year, island_side, island_side_habitat, site, habitat, site_habitat) %>%
  dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))
wide_site_level$site_habitat <- as.factor(wide_site_level$site_habitat)

site_portfolio_columns <- c("site_habitat", "portfolio_effect") # names of columns for dataframe
site_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(wide_site_level$site_habitat)), 
                    ncol = length(site_portfolio_columns))
             )
colnames(site_level_portfolio) <- site_portfolio_columns # change column 
site_level_portfolio$site_habitat <- unique(wide_site_level$site_habitat) # populate

for(i in 1:length(unique(wide_site_level$site_habitat)) ){ # for each  site
  # create a new dataframe for current site
  newData <- subset(wide_site_level, site_habitat == levels(wide_site_level$site_habitat)[i])
  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
  if(sum(current_matrix) == 0){ # if this was a 0 algae site
    site_level_portfolio[i, 2] <- NA # set portfolio effects to 0
  }else{ # otherwise
    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae

    site_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
  }

# merge together
diversity_stability_site <- merge(diversity_cv_site, lm_synchrony_site, by = "site_habitat")
diversity_stability_site <- merge(diversity_stability_site, site_level_portfolio, by = "site_habitat")

# write_csv(diversity_stability_site, here("data", "algae_diversity_stability_metrics_site.csv"))
